---
title: "Mitch Report"
author: "Antony Kaspi & Mark Ziemann"
output: html_document
theme: cosmo
---

date generated: `r Sys.Date()`

## Background

Mitch performs unidimensional and multidimensional gene set enrichment analysis. The
concept behind this dates to work by Cox and Mann 
(https://doi.org/10.1186/1471-2105-13-S16-S12). This implementation is suited to R based 
workflows of multi-omics datasets. This software was developed by Antony Kaspi and Mark 
Ziemann. Learn more about Mitch at the website: https://github.com/markziemann/Mitch

## Input profiles

Here is the first few lines of the input profile.

```{r,peek,echo=FALSE}
load(DATANAME)

# capture the dimensionality of the data                                               
d=ncol(res$input_profile)

# if working with >5 dimensions, then substitute the dimension (colnames) names with a number
if ( d>5 ) {

    mydims<-data.frame(seq_len(d))

    mydims$colnames<-attributes(res)$profile_dimensions

    colnames(mydims)<-c("dimension","contrast_name")

    print(kable( mydims, caption = "Profile dimensions" ))

    colnames(res$input_profile)<- paste("d",seq_len(d),sep="")

    colnames(res$ranked_profile)<- paste("d",seq_len(d),sep="")

    ss<-res$ranked_profile

}

head(res$input_profile)
```

Here are some metrics about the input data profile:
```{r, metrics, echo=FALSE}
if (d==1) {

    formatted<-t(as.data.frame(res$analysis_metrics[ c(1,2,3,4,5 ) ])) 

} else if (d==2) {

    unformatted<-t(as.data.frame(res$analysis_metrics[ c(2,3,4,5,11,12 )  ]))

    formatted<-unformatted

    formatted[1:4]<-as.character(round(as.numeric(unformatted[1:4]) , digits=0))

    formatted[5:6]<-as.character(round(as.numeric(unformatted[5:6]) , digits=5))

} else if (d>2) {

    formatted<-t(as.data.frame(res$analysis_metrics[ c(2,3,4,5 )  ]))

}

colnames(formatted)="Profile metrics"

kable( formatted, caption = "Profile metrics" )
```

Here is a plot of the input profiles. Note the dynamic ranges.
```{r, scatterplot, echo=FALSE,fig.height = 6, fig.width = 6.5, message=FALSE, warning=FALSE}
if ( d==1 ) {

    par(mfrow=c(2,1))

    hist(res$input_profile[,1],breaks=50,main="Distribution of DE scores",xlab=paste("DE score for ",colnames(res$input_profile)))

    plot(res$input_profile,xlab=paste("DE score for ",colnames(res$input_profile)),

    pch="|",frame.plot=FALSE)

    UPS=length(which(res$input_profile>0))

    DNS=length(which(res$input_profile<0))

    TOTAL=nrow(res$input_profile)

    mtext(paste(TOTAL,"genes in total,",UPS,"trending up-regulated,",DNS,"trending down-regulated"))

} else if ( d<3 ) {

    plot(res$input_profile, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.15), main="Input profiles"  )

} else {

    library("GGally")

    ggpairs_points_plot <- function(data ,mapping, ...){

        p <- ggplot(data = data, mapping = mapping) +
        geom_point(alpha=0.05) +
        geom_vline(xintercept=0,linetype="dashed") +
        geom_hline(yintercept=0,linetype="dashed")

    }

    p<-ggpairs(as.data.frame(res$input_profile), title="Scatterplot of all genes" , lower  = list(continuous = ggpairs_points_plot ))

    print( p +  theme_bw() )

}
```

Here is the contour plot of the profile including all detected genes.
```{r, contourplot, echo=FALSE,fig.height = 6, fig.width = 6.5,warning=FALSE, message=FALSE}
palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))

#Contour of all the data
ss<-res$ranked_profile

if (d==1) {

    message("Contour plot does not apply to unidimensional analysis.")

} else if (d==2) {

    xmin=min(ss[,1])

    xmax=max(ss[,1])

    ymin=min(ss[,2])

    ymax=max(ss[,2])

    ss<-res$ranked_profile

    k<-MASS:::kde2d(ss[,1],ss[,2])

    X_AXIS=paste("Rank in contrast",colnames(ss)[1])

    Y_AXIS=paste("Rank in contrast",colnames(ss)[2])

    filled.contour(k, xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    color=palette ,
    plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")
    title( main="Rank-rank plot of all genes",xlab=X_AXIS,ylab=Y_AXIS ) } )

} else if (d>2) {

    #pairs contour plot function
    ggpairs_func <- function(data, mapping, ...){

        p <- ggplot(data = data, mapping = mapping) +
        stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
        geom_vline(xintercept=0,linetype="dashed") +
        geom_hline(yintercept=0,linetype="dashed") +
        scale_fill_gradientn(colours=palette(25))

        p

    }

    #pairs contour plot
    p<-ggpairs(as.data.frame(ss), title="Contour plot of all genes after ranking" , 
    lower=list(continuous=ggpairs_func),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/100)))

    print( p + theme_bw() )

    #subset contour plot
    ggpairs_contour_limit_range <- function(data ,mapping, ...){

        p <- ggplot(data = data, mapping = mapping) +
        stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
        geom_vline(xintercept=0,linetype="dashed") +
        geom_hline(yintercept=0,linetype="dashed") +
        scale_fill_gradientn(colours=palette(25)) +
        scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
        scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )

        p

    }

    #subset points plot
    ggpairs_points_limit_range <- function(data ,mapping, ...){

        p <- ggplot(data = data, mapping = mapping) +
        geom_point(alpha=0.1) +
        geom_vline(xintercept=0,linetype="dashed") +
        geom_hline(yintercept=0,linetype="dashed") +
        scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
        scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )

        p

    }

}
```

## Input genesets
Here are some metrics about the gene sets used:
```{r,input_geneset_metrics1,results="asis",echo=FALSE}
ORIGINFILE=attributes(res$input_genesets)$originfile

cat(paste("GMT file of genesets:",ORIGINFILE,"<br>"))

unformatted<-t(as.data.frame(res$analysis_metrics[c(1,6,7)]))

formatted<-as.data.frame(as.character( unformatted[1:3]) )

rownames(formatted)=rownames(unformatted)

colnames(formatted)="Gene sets metrics"

kable( formatted , col.names = "Gene sets metrics" ,caption = "Gene sets metrics" )
```

```{r,input_geneset_metrics2,results="asis",echo=FALSE ,fig.height = 7, fig.width = 7 ,fig.show="all"}
par(mfrow=c(3,1))

geneset_counts<-res$analysis_metrics$geneset_counts

boxplot(geneset_counts$count,horizontal=TRUE,frame=FALSE,main="Gene set size",xlab="number of member genes included in profile")

hist(geneset_counts$count,100,xlab="geneset size",main="Histogram of geneset size")

hist(geneset_counts$count,100,xlim=c(0,500),xlab="geneset size",main="Trimmed histogram of geneset size")

if ( d==2 ) {

    uu=length(which(res$input_profile[,1]>0 & res$input_profile[,2]>0))

    ud=length(which(res$input_profile[,1]>0 & res$input_profile[,2]<0))

    dd=length(which(res$input_profile[,1]<0 & res$input_profile[,2]<0))

    du=length(which(res$input_profile[,1]<0 & res$input_profile[,2]>0))

    a<-as.data.frame(c(uu,ud,dd,du))

    rownames(a)=c("top-right","bottom-right","bottom-left","top-left")

    colnames(a)="a"

    par(mfrow=c(1,1))

    xx<-barplot(a$a,names.arg=rownames(a),main="number of genes in each quadrant")

    text(x = xx, y = a$a, label = a$a , pos = 1, cex = 1)

} else if (d>2) {

    if (d<6) {

        sig<-sign(ss)

        sector_count<-aggregate(seq_len(nrow(sig)) ~ ., sig, FUN = length)

        colnames(sector_count)[ncol(sector_count)]<-"Count"

        kable(sector_count ,caption = "Genes by sector" ,row.names=TRUE)

    }

}

```

```{r,input_geneset_metrics3,results="asis",echo=FALSE,message=FALSE,fig.height = 7, fig.width = 7}
if (d<2) {

    nsig=length(which(res$enrichment_result$p.adjustANOVA<0.05))

} else {

    nsig=length(which(res$enrichment_result$p.adjustMANOVA<0.05))
}

if (d==1) {

    cat("<h2>Differential pathway expression</h2><br>")

    par(mfrow=c(1,1))

    sig<-subset(res$enrichment_result,p.adjustANOVA<=0.05)

    plot(res$enrichment_result$s.dist,-log10(res$enrichment_result$pANOVA),
    xlab="s score",ylab="-log10(p-value)",
    main="volcano plot of gene set enrichments",pch=19,cex=0.8)

    points(sig$s.dist,-log10(sig$pANOVA),pch=19,cex=0.85,col="red")

    TOTAL=nrow(res$enrichment_result)

    SIG=nrow(sig)

    UP=length(which(sig$s.dist>0))

    DN=length(which(sig$s.dist<0))

    SUBHEADER=paste(TOTAL,"gene sets in total,",UP,"upregulated and ",DN,"downregulated (FDR<=0.05)")

    mtext(SUBHEADER)
}

if (d==2) {

    cat("<h2>Gene sets by quadrant</h2><br>")

    cat(paste("Number of significant gene sets (FDR<0.05)=", res$analysis_metrics$num_sets_significant,"<br>" ))

    a<-res$analysis_metrics[14]

    a<-as.data.frame(as.numeric(unlist(strsplit(as.character(a),','))),stringsAsFactors=FALSE)

    rownames(a)=c("top-right","bottom-right","bottom-left","top-left")

    colnames(a)="a"

    xx<-barplot(a$a,names.arg=rownames(a),main="number of genesets FDR<0.05")

    text(x = xx, y = a$a, label = a$a , pos = 1, cex = 1)
} 

if ( nsig > 0 ) {

    if ( d>2 ) {

        if ( d<6 ) {

            cat(paste("Number of significant gene sets (FDR<0.05)=", res$analysis_metrics$num_sets_significant,"<br>" ))

            cat("<h2>Gene sets by sector</h2><br>")

            sig<-sign(res$enrichment_result[which(res$enrichment_result$p.adjustMANOVA<0.05),4:(4+d-1)])

            sector_count<-aggregate(seq_len(nrow(sig)) ~ ., sig, FUN = length)

            colnames(sector_count)[ncol(sector_count)]<-"Count"

            kable(sector_count ,caption = "Gene sets by sector" ,row.names=TRUE)
        }
    }
}

```

## Enrichment scatterplot with taucharts
```{r,tauchart1d,results="asis",echo=FALSE, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}

if (d==1) {

    #1D tauchart
    library("taucharts")
    library(dplyr)
    library(gtools)

    numsets=nrow(subset(res$enrichment_result,p.adjustANOVA<0.05))

    p=NULL

    if (numsets==0){

        message("No significant enrichments found.")

    } else {

        # volcano tauchart
        myres<-subset(res$enrichment_result,p.adjustANOVA<0.05)

        myres$significance<--log10(myres$pANOVA)

        myres$set<-gsub("_"," ",as.character(myres$set))

        myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

        colnames(myres)=gsub("\\.","_",colnames(myres))

        my_x=colnames(myres)[4]

        my_y=colnames(myres)[6]

        p<-tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

        cat("Significance is calculated by -log10(p-value). All points shown are FDR<0.05.<br>")

    }

    p
}

if (d==1) {

    resrows=length(res$detailed_sets)

    if (resrows>1) {

        myres<-head(res$enrichment_result,resrows)

        myres$significance<--log10(myres$pANOVA)

        myres$set<-gsub("_"," ",as.character(myres$set))

        myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

        colnames(myres)=gsub("\\.","_",colnames(myres))

        my_x=colnames(myres)[4]

        my_y=colnames(myres)[6]

        p<-tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

        cat("Significance is calculated by -log10(p-value). Top N sets shown irrespective of FDR.<br>")

    }

    p

}

```{r,tauchart2,results="asis",echo=FALSE, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}

tauscatter<-function(i){

    my_x=plan[i,1]

    my_y=plan[i,2]

    colnames(myres)=gsub("\\.","_",colnames(myres))

    my_x=colnames(myres)[3+my_x]

    my_y=colnames(myres)[3+my_y]

    myres$set<-gsub("_"," ",as.character(myres$set))

    tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

}


if (d!=1) {
    library("taucharts")
    library(dplyr)
    library(gtools)

    cat("All sets with FDR<0.05. Try hovering over the points.<br>")

    numsets=nrow(subset(res$enrichment_result,p.adjustMANOVA<0.05))

    p=NULL

    if (numsets<1){

        message("No significant enrichments found.")

    } else {

        myres<-subset(res$enrichment_result,p.adjustMANOVA<0.05)

        if ( d<3 ) {

            myres<-myres[,c(1,which(names(myres) %in% "p.adjustMANOVA"),2, 4:(4+d-1) , which(names(myres) %in% "s.dist") )]

            myres$set<-gsub("_"," ",as.character(myres$set))

            colnames(myres)=gsub("\\.","_",colnames(myres))

            my_x=colnames(myres)[4]

            my_y=colnames(myres)[5]

            p<-tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

        } else {

            plan<-combinations(n = d, r = 2, v = seq_len(d), repeats.allowed = FALSE)

            myres<-myres[,c(1, which(names(myres) %in% "p.adjustMANOVA") ,2,4:(4+d-1), which(names(myres) %in% "s.dist") ), which(names(myres) %in% "SD")]

            p<-list()

            p<-lapply( seq_len(nrow(plan)) , tauscatter)

        }

    }

    htmltools::tagList(p)

}

if (d!=1) {

    cat("Top N sets irrespective of FDR. Try hovering over the points.<br>")

    resrows=length(res$detailed_sets)

    p=NULL

    if (resrows<1){

        message("No results found.")

    } else {

        myres<-head(res$enrichment_result,resrows)

        if ( d<3 ) {

            myres<-myres[,c(1,which(names(myres) %in% "p.adjustMANOVA"),2, 4:(4+d-1) , which(names(myres) %in% "s.dist") )]

            myres$set<-gsub("_"," ",as.character(myres$set))

            colnames(myres)=gsub("\\.","_",colnames(myres))

            my_x=colnames(myres)[4]

            my_y=colnames(myres)[5]

            p<-tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

        } else {

            plan<-combinations(n = d, r = 2, v = seq_len(d), repeats.allowed = FALSE)

            myres<-myres[,c(1, which(names(myres) %in% "p.adjustMANOVA") ,2,4:(4+d-1), which(names(myres) %in% "s.dist") ), which(names(myres) %in% "SD")]

            p<-list()

            p<-lapply( seq_len(nrow(plan)) , tauscatter)

        }

    }

    htmltools::tagList(p)

}

```

```{r,heatmap,results="asis",echo=FALSE, fig.height = 10, fig.width = 7 ,fig.show="all", message=FALSE}
if (d!=1) {

    if (resrows>2) {

        cat("<h2> A heatmap of S values for top results</h2><br>")

        library(gplots)

        hmapx<-head( res$enrichment_result[,4:(4+d-1)] ,resrows)

        rownames(hmapx)<-head(res$enrichment_result$set,resrows)

        colnames(hmapx)<-gsub("^s.","",colnames(hmapx))

        my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 25)

        if(resrows<25){

            CEXROW=1

        } else if (resrows<51) {

            CEXROW=0.8

        } else if (resrows<76) {

            CEXROW=0.6

        } else if (resrows<101) {

            CEXROW=0.4

        } else {

            CEXROW=0.25

        }

        heatmap.2(as.matrix(hmapx),scale="none",margin=c(10, 25),cexRow=CEXROW,trace="none",cexCol=1,col=my_palette)

    }

}
```

```{r,effectsize,results="asis",echo=FALSE, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}
if (d!=1) {

    cat('<h2> A plot of effect size versus significance</h2><br>')

    cat('Significance is the -log2(p.adjustMANOVA) and effect size is the s.dist which is the hypotenuse of the s scores.<br>')

    myres<-res$enrichment_result

    myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

    myres$significance<--log2(myres$p.adjustMANOVA)

    myres$set<-gsub("_"," ",as.character(myres$set))

    colnames(myres)=gsub("\\.","_",colnames(myres))

    my_x=colnames(myres)[ncol(myres)-2]

    my_y=colnames(myres)[ncol(myres)]

    tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

}
```

## Results table
```{r,results_table,results="asis",echo=FALSE}
if ( d==1 ) {

    resrows=length(res$detailed_sets)

    myres<-head(res$enrichment_result,resrows)

    myres[,c(3:ncol(myres))]<-signif(myres[,c(3:ncol(myres))],3)

} else if (d>1) {

    resrows=length(res$detailed_sets)

    myres<-head(res$enrichment_result,resrows)

    myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist") ,4:((d*2)+3)), which(names(myres) %in% "SD")]

    myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
}

formatted<-myres

formatted[,1]<-gsub("_"," ",myres[,1])

kable( formatted , col.names = colnames(formatted) , row.names=FALSE, caption = cat(paste("Top N=",resrows,"gene sets")) ,digits=100)

cat("<hr><br>")
```

## Results (complete table)
```{r,results_table_complete,results="asis",echo=FALSE}
myres<-res$enrichment_result

if ( d==1 ) {

    myres[,c(3:ncol(myres))]<-signif(myres[,c(3:ncol(myres))],3)

    formatted<-myres

    formatted[,1]<-gsub("_"," ",myres[,1])

} else if (d>1) {

    myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist") ,4:((d*2)+3)), which(names(myres) %in% "SD")]

    myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

    formatted<-myres

    formatted[,1]<-gsub("_"," ",myres[,1])

    formatted[,3:ncol(myres)]<-format(myres[,3:ncol(myres)],digits=3)

}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )

cat(HEADER)

kable(formatted, col.names=colnames(formatted),row.names=FALSE,caption = "Complete results",digits=100)

cat("<br></p></details>")

cat("<hr><br>")
```

```{r,detailed_geneset_reports1d,results="asis",echo=FALSE,fig.height=6, fig.width=6 ,out.width = '80%',comment=NA, message=FALSE}
if (d==1) {

    cat("<h2>Detailed Gene set reports</h2><br>")

    cat('<br>')

    cat('\n')

    library("beeswarm")

    ss<-res$ranked_profile

    for ( i in seq_len(resrows) ) {

        mydat = NULL

        mydat<-t(myres[i,])

        cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))

        print(kable(mydat,format='markdown',caption=as.character(t(myres[i,1]))) )

        cat('\n')

        cat("<br>")

        # plots
        sss<-res$detailed_sets[[i]]

        set<-names(res$detailed_sets[i])

        size<-length(sss)

        par(mfrow=c(3,1))

        beeswarm(sss,vertical = FALSE,cex=0.75,xlim=c(min(ss),
        max(ss)),col="darkgray",pch=19,main=set,cex.main=1.5,
        xlab=paste("ranked DE score in:",colnames(ss)))

        mtext("beeswarm plot",cex=0.8)

        hist(sss,xlim=c(min(ss),max(ss)),breaks=15,col="darkgray",main=NULL,
        border="black",xlab=paste("ranked DE score in:",colnames(ss)))

        mtext("histogram",cex=0.8)

        plot(sss,rep(1,length(sss)),type="n",xlim=c(min(ss),max(ss)),
        frame=FALSE,axes=FALSE,ylab="",
        xlab=paste("ranked DE score in:",colnames(ss)))

        rug(sss, ticksize = 0.9)

        axis(1)

        mtext("rugplot",cex=0.8)

        cat("<br>")

        # top gene list
        setSign=sign(res$enrichment_result[i,4])

        if (setSign==-1) {

            tops<-as.data.frame(res$detailed_sets[[i]][order(res$detailed_sets[[i]])])

        } else if (setSign==1) {

            tops<-as.data.frame(res$detailed_sets[[i]][order(-res$detailed_sets[[i]])])

        }

        cat("Top enriched genes\\n")

        tops$GeneID<-rownames(tops)

        tops<-tops[,c(2,1)]

        colnames(tops)=c("GeneID","Gene Rank")

        cat("<br>")

        print(kable(head(tops,n=20L),col.names=colnames(tops),format="markdown",row.names=FALSE,caption="Top 20 genes",digits=100))

        cat('\n')

        HEADER=paste("<br><details><summary><b>","Click HERE to show all gene set members","</b></summary><br><p>",sep=" " )

        cat(HEADER)

        print(kable(tops,col.names=colnames(tops),format="markdown",row.names=FALSE,caption="All member genes",digits=100))

        cat('\n')

        cat("<br></p></details>")

        cat("<br><hr>")

    }

}
```



```{r,detailed_geneset_reports2d,results="asis",echo=FALSE,fig.height=5, fig.width=6 ,out.width = '80%',comment=NA, message=FALSE}

twodimplot<-function(i) {

    ll<-res$enrichment_result[i,]

    size<-ll$setSize

    sss<-res$detailed_sets[[i]]

    k<-MASS:::kde2d(sss[,1],sss[,2])

    filled.contour( k, color = palette, xlim=c(xmin,xmax),ylim=c(ymin,ymax),

    plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")

    title( main=ll$set , xlab=X_AXIS,ylab=Y_AXIS )})

    plot(sss, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.2),
    main=ll$set , xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    xlab=X_AXIS,ylab=Y_AXIS)

    abline(v=0,h=0,lty=2,lwd=2,col="blue")

    sss_long<-melt(sss)

    p<-ggplot(ss_long,aes(Var2,value)) +
    geom_violin(data=ss_long,fill = "grey", colour = "grey") +
    geom_boxplot(data=ss_long,width=0.9,fill="grey",outlier.shape = NA) +
    geom_violin(data=sss_long,fill = "black", colour = "black") +
    geom_boxplot(data=sss_long,width=0.1,outlier.shape = NA) +
    labs(y = "Position in rank",title = ll[,1] )

    print( p + theme_bw() +
    theme( axis.text=element_text(size=14),
    axis.title=element_text(size=15),
    plot.title = element_text(size = 14)))
}


# subset contour plot
ggpairs_contour_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
    stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
    geom_vline(xintercept=0,linetype="dashed") +
    geom_hline(yintercept=0,linetype="dashed") +
    scale_fill_gradientn(colours=palette(25)) +
    scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
    scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
}


# subset points plot
ggpairs_points_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
    geom_point(alpha=0.1) +
    geom_vline(xintercept=0,linetype="dashed") +
    geom_hline(yintercept=0,linetype="dashed") +
    scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
    scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
}



ndplot<-function(i) {

    ll<-res$enrichment_result[i,]

    size<-ll$setSize

    ss<-res$ranked_profile

    sss<-res$detailed_sets[[i]]

    if ( d>5 ) {

        colnames(sss)<- paste("d",seq_len(d),sep="")

    }

    p<-ggpairs(as.data.frame(sss), title=ll[,1], lower=list(continuous=ggpairs_contour_limit_range),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)) )

    print( p + theme_bw() )

    p<-ggpairs(as.data.frame(sss), title=ll[,1], lower= list(continuous = ggpairs_points_limit_range ),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)))

    print( p + theme_bw() )

    sss_long<-melt(sss)

    p<-ggplot(ss_long,aes(Var2,value)) +
    geom_violin(data=ss_long,fill = "grey", colour = "grey") +
    geom_boxplot(data=ss_long,width=0.9,fill="grey",outlier.shape = NA) +
    geom_violin(data=sss_long,fill = "black", colour = "black") +
    geom_boxplot(data=sss_long,width=0.1,outlier.shape = NA) +
    labs(y = "Position in rank",title = ll[,1] )

    print( p + theme_bw() +
    theme( axis.text=element_text(size=14),
    axis.title=element_text(size=15),
    plot.title = element_text(size = 14)))

}



topgenelist<-function(i) {
    ss<-res$ranked_profile

    sss<-res$detailed_sets[[i]]

    if ( d>5 ) {

        colnames(sss)<- paste("d",seq_len(d),sep="")

    }

    tl=bl=tr=br=0

    myres<-head(res$enrichment_result,resrows)

    myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist") ,4:((d*2)+3)) , which(names(myres) %in% "s.dist")]

    myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

    #select 2 strongest columns to highlight genes
    if( d>2 ) {

        cols<-(ncol(myres)-(2*d)+1):(ncol(myres)-(2*d)+d)

        COLS<-order(-abs(myres[i,cols]))[1:2]

        sss<-sss[,COLS]

        cols<-cols[COLS]

        mysx=signif(myres[i,cols[1]],3)

        mysy=signif(myres[i,cols[2]],3)

    } else {

        cols=c(ncol(myres)-3,ncol(myres)-2)

        mysx=signif(myres[i,cols[1]],3)

        mysy=signif(myres[i,cols[2]],3)

    }

    if ( mysy>0 ) { tl=tl+1 ; tr=tr+1  } else { bl=bl+1 ; br=br+1 }

    if ( mysx>0 ) { tr=tr+1 ; br=br+1  } else { tl=tl+1 ; bl=bl+1 }

    if (bl==2) {

        myquad<-sss[which(sss[,1]<0 & sss[,2]<0),,drop=FALSE]

        topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),,drop=FALSE]
    }

    if (tr==2) {

        myquad<-sss[which(sss[,1]>0 & sss[,2]>0),,drop=FALSE]

        topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),,drop=FALSE]

    }

    if (br==2) {

        myquad<-sss[which(sss[,1]>0 & sss[,2]<0),,drop=FALSE]

        topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),,drop=FALSE]

    }

    if (tl==2) {

        myquad<-sss[which(sss[,1]<0 & sss[,2]>0),,drop=FALSE]

        topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),,drop=FALSE]

    }

    cat("<br>")

    topgenes<-as.data.frame(topgenes)

    topgenes$Gene<-as.character(rownames(topgenes))

    topgenes<-topgenes[,c(3,1,2)]

    print(kable(head(topgenes,n=20L),col.names=colnames(topgenes),format="markdown",row.names=FALSE,caption="Top 20 genes",digits=100))

    cat('\n')

    HEADER=paste("<br><details><summary><b>","Click HERE to show all gene set members","</b></summary><br><p>",sep=" " )

    cat(HEADER)

    sss<-res$detailed_sets[[i]]

    print(kable(sss,format="markdown",row.names=TRUE,caption="All member genes",digits=100))

    cat('\n')

    cat("<br></p></details>")

    cat("<br><hr>")

}


#functions end here

if (d!=1) {

    cat("<h2> Detailed Gene set reports</h2><br>")

    library("reshape2")

    library("ggplot2")

    library("GGally")

    resrows=length(res$detailed_sets)

    ss<-res$ranked_profile

    myres<-head(res$enrichment_result,resrows)

    myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist") ,4:((d*2)+3)), which(names(myres) %in% "SD")]

    myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

    palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))

    ss_long<-melt(ss)

    if ( d<3 ) {

        xmin=min(ss[,1])

        xmax=max(ss[,1])

        ymin=min(ss[,2])

        ymax=max(ss[,2])

            for ( i in seq_len(resrows )) {

                mydat = NULL                                                    

                mydat$metrics<-names(myres[2:ncol(myres)])

                mydat$values<-unname(t(myres[i,2:ncol(myres)]))

                mydat<-as.data.frame(cbind(mydat$metrics,mydat$values))

                colnames(mydat)<-c("metric","value")

                cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))

                print(kable(mydat,format='markdown',caption=as.character(t(myres[i,1]))) )

                cat('\n')

                cat("<br>")

                twodimplot(i)

                cat("<br>")

                topgenelist(i)

                cat("<hr><br>")

            }

        } else {

        for ( i in seq_len(resrows )) {

            mydat = NULL

            mydat$metrics<-names(myres[2:ncol(myres)])

            mydat$values<-unname(t(myres[i,2:ncol(myres)]))

            mydat<-as.data.frame(cbind(mydat$metrics,mydat$values))

            colnames(mydat)<-c("metric","value")

            cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))

            print(kable(mydat,format='markdown',digits=5,caption=as.character(t(myres[i,1]))))

            cat('\n')

            cat("<br>")

            ndplot(i)

            cat("<br>")

            topgenelist(i)

            cat("<hr><br>")

        }

    }

}
```

Here is the session info with all the versions of packages used.

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
library(pkgload)
sessionInfo()
```

END of report
